---
title: "Linear Model and Contrast Matrix"
author: "Marta Ibáñez Lligoña"
date: "5/29/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Description of the project
Cellular therapy is proposed for tendinopathy treatment. Bone marrow- (BM-MSC) and adipose tissue- (ASC-) derived mesenchymal stromal cells are candidate populations for such a therapy.
We used microarrays to evaluate the basal gene expression in human BM-MSCs and ASCs and to create list of differentially expressed genes between analyzed groups.

## Loading targets

```{r}
require(Biobase)
require (limma)
workingDir <-getwd()
targets <-read.csv(file=file.path(workingDir,"targets.csv"), header = TRUE, sep=";")

```

## Linear Model

```{r}
design <- matrix(c(1,1,1,1,1,0,0,0,0,0,0,0,0,0,0,0,1,1,1,1,1,1), nrow = 11, byrow = F)
design2 <-model.matrix(~ 0+targets$Group)
colnames(design)<- colnames(design2)<-c("ASC", "BM_BSC")
rownames(design) <- rownames(design2)<- targets$ShortName
print(design); print(design2)
```

## Contrast Matrix
```{r}
cont.matrix <- makeContrasts (
  ASC_BMBSC = ASC-BM_BSC,
  levels=design)
```