---
title: "MA analysis"
author: "Marta Ibáñez Lligoña"
date: "5/27/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Looking for the data

First I went to GEO and looked for the dataset that I needed to do the exercise, I downloaded the file. 
I went through the folder and extracted all files so I didn't have anY zip files when working with it.

```{r}
require(Biobase)
workingDir <-getwd()
dataDir <- file.path(workingDir, "data")
resultsDir <- file.path(workingDir, "results")
```

# Loading data

## Targets

```{r}
targets <-read.csv(file=file.path(dataDir,"targets.csv"), header = TRUE, sep=";")
targets
```

## CEL Files

```{r}

CELfiles <- list.celfiles(file.path(dataDir)) #different cel files in data
CELfiles
rawData <- read.celfiles(file.path(dataDir,CELfiles)) #cel files in another way

# Define some variables for plots

```{r}
sampleNames<- as.character(targets$ShortName)
sampleNames
sampleColor <- as.character(targets$Color)
sampleColor
```

# BOXPLOT
```{r}
boxplot(rawData, which="all",las=2, main="Intensity distribution of RAW data", 
        cex.axis=0.6, names = sampleNames, col=sampleColor)
```


# HIERARCHICAL CLUSTERING

```{r}
clust.euclid.average <- hclust(dist(t(exprs(rawData))),method="average")
plot(clust.euclid.average, labels=sampleNames, main="Hierarchical clustering of RawData", 
     cex=0.7,  hang=-1)
```

# PRINCIPAL COMPONENT ANALYSIS

```{r}
plotPCA <- function ( X, labels=NULL, colors=NULL, dataDesc="", scale=FALSE, formapunts=NULL, myCex=0.8,...)
{
  pcX<-prcomp(t(X), scale=scale) # o prcomp(t(X))
  loads<- round(pcX$sdev^2/sum(pcX$sdev^2)*100,1)
  xlab<-c(paste("PC1",loads[1],"%"))
  ylab<-c(paste("PC2",loads[2],"%"))
  if (is.null(colors)) colors=1
  plot(pcX$x[,1:2],xlab=xlab,ylab=ylab, col=colors, pch=formapunts, 
       xlim=c(min(pcX$x[,1])-100000, max(pcX$x[,1])+100000),ylim=c(min(pcX$x[,2])-100000, max(pcX$x[,2])+100000))
  text(pcX$x[,1],pcX$x[,2], labels, pos=3, cex=myCex)
  title(paste("Plot of first 2 PCs for expressions in", dataDesc, sep=" "), cex=0.8)
}

plotPCA(exprs(rawData), labels=sampleNames, dataDesc="raw data", colors=sampleColor,
        formapunts=c(rep(16,4),rep(17,4)), myCex=0.6)
```


## Save to file
```{r}
pdf(file.path(resultsDir, "QCPlots_Raw.pdf"))
boxplot(rawData, which="all",las=2, main="Intensity distribution of RAW data", 
        cex.axis=0.6, col=sampleColor, names=sampleNames)
plot(clust.euclid.average, labels=sampleNames, main="Hierarchical clustering of samples of RawData", 
     cex=0.7,  hang=-1)
plotPCA(exprs(rawData), labels=sampleNames, dataDesc="raw data", colors=sampleColor,
        formapunts=c(rep(16,4),rep(17,4)), myCex=0.6)
dev.off()
```
# DATA NORMALIZATION

```{r}
eset<-rma(rawData)

write.exprs(eset, file.path(resultsDir, "NormData.txt"))
```

## Quality control of arrays: Normalized data

### Boxplot

```{r}
boxplot(eset, las=2, main="Intensity distribution of Normalized data", cex.axis=0.6, 
        col=sampleColor, names=sampleNames)
```

### Hierarchical clustering

```{r}
clust.euclid.average <- hclust(dist(t(exprs(eset))),method="average")
plot(clust.euclid.average, labels=sampleNames, main="Hierarchical clustering of NormData", 
     cex=0.7,  hang=-1)
```

## Principal Component Analysis

```{r}
plotPCA <- function ( X, labels=NULL, colors=NULL, dataDesc="", scale=FALSE, formapunts=NULL, myCex=0.8,...)
{
  pcX<-prcomp(t(X), scale=scale) # o prcomp(t(X))
  loads<- round(pcX$sdev^2/sum(pcX$sdev^2)*100,1)
  xlab<-c(paste("PC1",loads[1],"%"))
  ylab<-c(paste("PC2",loads[2],"%"))
  if (is.null(colors)) colors=1
  plot(pcX$x[,1:2],xlab=xlab,ylab=ylab, col=colors, pch=formapunts, 
       xlim=c(min(pcX$x[,1])-10, max(pcX$x[,1])+10),ylim=c(min(pcX$x[,2])-10, max(pcX$x[,2])+10))
  text(pcX$x[,1],pcX$x[,2], labels, pos=3, cex=myCex)
  title(paste("Plot of first 2 PCs for expressions in", dataDesc, sep=" "), cex=0.8)
}
plotPCA(exprs(eset), labels=sampleNames, dataDesc="NormData", colors=sampleColor,
        formapunts=c(rep(16,4),rep(17,4)), myCex=0.6)
```

## Saving to file

```{r}
pdf(file.path(resultsDir, "QCPlots_Norm.pdf"))
boxplot(eset, las=2, main="Intensity distribution of Normalized data", cex.axis=0.6, 
        col=sampleColor, names=sampleNames)
plot(clust.euclid.average, labels=sampleNames, main="Hierarchical clustering of NormData", 
     cex=0.7,  hang=-1)
plotPCA(exprs(eset), labels=sampleNames, dataDesc="selected samples", colors=sampleColor,
        formapunts=c(rep(16,4),rep(17,4)), myCex=0.6)
dev.off()
```

# Array Quality Metrics

# LINEAR MODEL AND CONTRAST MATRIX
## Linear model matrix
```{r}
design <- matrix(c(1,1,1,1,1,0,0,0,0,0,0,0,0,0,0,0,1,1,1,1,1,1), nrow = 11, byrow = F)
design2 <-model.matrix(~ 0+targets$Group)
colnames(design)<- colnames(design2)<-c("ASC", "BM_BSC")
rownames(design) <- rownames(design2)<- targets$ShortName
print(design); print(design2)
```

## Contrast matrix
```{r}
cont.matrix <- makeContrasts (
  ASC_BMBSC = ASC-BM_BSC,
  levels=design)
```
